---
title: "BenchmarkeR"
author: "Alicia Brown"
output: 
  flexdashboard::flex_dashboard:
    theme: cerulean
    orientation: rows
    social: menu
    source_code: embed
    favicon: favicon.png
runtime: shiny
---

```{r setup, include=FALSE}
library(dplyr)
library(flexdashboard)
library(ggplot2)
library(httr)
library(jsonlite)
library(lubridate)
library(readr)
library(shiny)
library(stringr)
library(tidyr)
require(wesanderson)

api_mode = T
```

Inputs {.sidebar data-width=300}
-----------------------------------------------------------------------

```{r input}
textAreaInput("domains", label = "Domains:", value = "data.seattle.gov,data.mesaaz.gov,www.dallasopendata.com", placeholder = "Domain", resize = "vertical")

sliderInput("description_weight", label = "Meta Description weight", min = 0, max = 1, value = 1, step = .25, round = FALSE,
  format = NULL, locale = NULL, ticks = T, animate = FALSE, dragRange = TRUE)
sliderInput("category_weight", label = "Meta Category weight", min = 0, max = 1, value = 1, step = .25, round = FALSE,
  format = NULL, locale = NULL, ticks = T, animate = FALSE, dragRange = TRUE)
sliderInput("keywords_weight", label = "Meta Keywords weight", min = 0, max = 1, value = 1, step = .25, round = FALSE,
  format = NULL, locale = NULL, ticks = T, animate = FALSE, dragRange = TRUE)
sliderInput("source_weight", label = "Meta Source weight", min = 0, max = 1, value = 1, step = .25, round = FALSE,
  format = NULL, locale = NULL, ticks = T, animate = FALSE, dragRange = TRUE)
sliderInput("custom_weight", label = "Meta Custom Metadata weight", min = 0, max = 1, value = 1, step = .25, round = FALSE,
  format = NULL, locale = NULL, ticks = T, animate = FALSE, dragRange = TRUE)

# download_count not available to pass to order parameter
selectInput("rank", label = "Sort datasets by:",
            choices = list(`Views Last Month` = "page_views_last_month",
           `Views Last Week` = "page_views_last_week",
           `Total Views` = "page_views_total"), selected = "page_views_last_month")

renderText("Powered by https://socratadiscovery.docs.apiary.io.")
```

```{r primary_url}
primary_domain <- reactive({
  first_domain <- str_split(input$domains,",",simplify = TRUE)
  first_domain <- first_domain[[1]]
  first_domain
})
```

```{r disco_url}
discoURL <- reactive({
  
  #10,000 results seems to be enough for now
  domains <- input$domains
  
  discovery_url <- paste0("http://api.us.socrata.com/api/catalog/v1?published=true&only=dataset"
                          ,"&limit=10000"
                          ,"&order=",input$rank
                          ,"&domains=",input$domains)
  discovery_url
})
```

```{r disco_data}
# Reactive data available for all shiny modules
discoData <- reactive({
  
  if(api_mode == F){
    static_file <- 'data/total_response.json'
    if(input$rank == 'page_views_last_week'){
      static_file <- 'data/last_7_days_response.json'
    }
    if(input$rank == 'page_views_last_month'){
      static_file <- 'data/last_30_days_response.json'
    }
    
    datasets <- read_json(static_file,simplifyDataFrame = TRUE)
    
  }else{
    datasets <- fromJSON(discoURL(),simplifyDataFrame = TRUE)
  }
  
  # debug
  # datasets <- fromJSON('http://api.us.socrata.com/api/catalog/v1?published=true&only=dataset&limit=10000&order=page_views_last_month&domains=www.dallasopendata.com',simplifyDataFrame = TRUE)
  
  resource <- datasets$results$resource
  
  if (is.null(resource) == FALSE){
    raw_data <- resource %>%
      select(id,name,description,attribution,updatedAt,createdAt,download_count)
    
    # page view stats/update/create
    page_views <- resource$page_views
    last_updated <- resource$updatedAt
    created <- resource$createdAt
    
    # meta
    meta <- datasets$results$metadata
    
    # link
    link <- datasets$results$link
    
    # domain
    domain <- meta$domain
    
    # domain category
    classification <- datasets$results$classification %>%
      select(domain_category, domain_tags, domain_metadata)
    
    data <- cbind(domain, raw_data, page_views, last_updated, created, link, classification)
    
    # set selected rank variable for plots to adjust y variable easily
    if(input$rank == 'page_views_last_month'){
     data <- data %>%
       mutate(measure_value = page_views_last_month)
    }
    if(input$rank == 'page_views_last_week'){
      data <- data %>%
       mutate(measure_value = page_views_last_week)
    }
    if(input$rank == 'page_views_total'){
      data <- data %>%
       mutate(measure_value = page_views_total)
    }
    
    # capitalize x variables, calculate meta score and replace NAs with 0 in download_count
    data <- data %>%
      mutate(download_count = replace_na(download_count, 0)) %>%
      mutate(days_last_updated = (now() - ymd_hms(last_updated))/ddays(1)) %>%
      mutate(has_description = ifelse(is.na(description), 0, 1 * input$description_weight)) %>%
      mutate(has_category = ifelse(is.na(domain_category), 0, 1 * input$category_weight)) %>%
      mutate(has_source = ifelse(is.na(attribution), 0, 1 * input$source_weight)) %>%
      mutate(has_keywords = ifelse(is.na(domain_tags), 0, 1 * input$keywords_weight)) %>%
      # empty list() = 6 characters
      mutate(has_custom = ifelse(nchar(domain_metadata) > 6, 1 * input$custom_weight, 0)) %>%
      mutate(`Metadata Score` = (has_description + has_category + has_source + has_keywords + has_custom)/(input$description_weight + input$category_weight + input$source_weight + input$keywords_weight + input$custom_weight) * 100)

    data
    }else{
      NULL
    }
})
```

Row
-----------------------------------------------------------------------

### Primary Domain {.value-box}

```{r primary_domain}
renderValueBox({
  domain <- primary_domain()
  
  valueBox(domain, 
           icon = "fa-city")
})
```

Row
-----------------------------------------------------------------------

### Datasets {.value-box}

```{r domain_dataset_count}
renderValueBox({
  data_view <- discoData()
  
  data_view <- data_view %>%
    filter(domain == primary_domain())
    
  assets_count <- nrow(data_view)
  
  valueBox(assets_count, 
           icon = "fa-area-chart",
           color = "info")
  
})
```

### Median Last Update (days) {.value-box}

```{r domain_days_last_update, eval = T}
renderValueBox({
  
  data_view <- discoData() %>%
    filter(domain == primary_domain())
    
  median_days <- median(data_view$days_last_updated)
  median_days <- format(median_days, digits = 1)
  
  valueBox(median_days, 
           icon = "fa-calendar",
           color = ifelse(median_days > 40, "warning", "success"))
  
})
```

### Average Meta Score {.gauge}

```{r primary_scores}
renderGauge({
  data_view <- discoData()
  
  data_view <- data_view %>%
    filter(domain == primary_domain())
  
  meta_score <- mean(data_view$`Metadata Score`)
  
  meta_score <- format(meta_score, digits = 1)
  
  gauge(meta_score, min = 0, max = 100, symbol = '%', gaugeSectors(
    success = c(90, 100), warning = c(50, 89), danger = c(0, 49)
  ))
  
})
```

Row {.tabset}
-----------------------------------------------------------------------

### Domain Metadata Scores

```{r domain_scores, fig.width=18, fig.height=8}
renderPlot({
  
  data_view <- discoData() %>%
    group_by(domain) %>%
    summarize(`Metadata Score` = mean(`Metadata Score`))
  
  if(length(data_view) > 0){
    # set sort order (todo: check out forcats package)
    data_view$domain <- factor(data_view$domain, levels=unique(data_view$domain[order(data_view$`Metadata Score`)]))
    
    # view Metadata score by domain - todo: add label to bar with score %
    ggplot(data_view) +
    geom_bar(mapping = aes(x = domain, y = `Metadata Score`, fill=domain), na.rm = TRUE, stat = "identity") +
    scale_fill_manual(values = wes_palette("Zissou1")) + 
    coord_flip() +
    labs(x = '', y = '', 
        title=paste0('Metadata Score by ',str_to_title(str_replace_all(input$rank,'_',' ')))) +
theme(text = element_text(size=15), plot.title = element_text(size=15, hjust=0.5))
  }
})
```

### Dataset Scores & Stats
    
```{r fig.width=18, fig.height=8}

renderTable({
  
  data_view <- discoData()
   
  if(length(data_view) > 0){
    data_view <- data_view %>%
      
    # format numbers with commas so easier to read in table view
    mutate(download_count = format(download_count, big.mark=","),
           page_views_last_month = format(page_views_last_month, big.mark=","),
           page_views_last_week = format(page_views_last_week, big.mark=","),
           page_views_total = format(page_views_total, big.mark=","),
           # create href tags
           link = paste0('<a href="',link,'" target="_blank">',name,'</a>'),
           domain = paste0('<a href="https://',domain,'/browse" target="_blank">',domain,'</a>')) %>%
    rename(`Views Last Month` = page_views_last_month,
           `Views Last Week` = page_views_last_week,
           `Views Total` = page_views_total,
           `Downloads Total` = download_count,
           `Last Updated (days)` = days_last_updated) %>%
    select(link, domain, `Metadata Score`, `Last Updated (days)`,`Views Last Week`,`Views Last Month`,`Views Total`,`Downloads Total`)

  data_view
  }
  else{
    paste0('No results found for <a href="',discoURL(),'" target="_blank">',discoURL(),'</a>')
  }
}
# allow html hyperlinks
, sanitize.text.function = function(x) x)
```

### Individual Metadata Scores
    
```{r fig.width=18, fig.height=8}

renderTable({
  
  data_view <- discoData()
   
  if(length(data_view) > 0){
    data_view <- data_view %>%
      
    # format numbers with commas so easier to read in table view
    mutate(download_count = format(download_count, big.mark=","),
           page_views_last_month = format(page_views_last_month, big.mark=","),
           page_views_last_week = format(page_views_last_week, big.mark=","),
           page_views_total = format(page_views_total, big.mark=","),
           # create href tags
           link = paste0('<a href="',link,'" target="_blank">',name,'</a>'),
           domain = paste0('<a href="https://',domain,'/browse" target="_blank">',domain,'</a>')) %>%
    rename(`Has Description` = has_description,
           `Has Category` = has_category,
           `Has Keywords` = has_keywords,
           `Has Source` = has_source,
           `Has Custom` = has_custom) %>%
    select(link, domain,`Metadata Score`,`Has Description`,`Has Category`,`Has Keywords`,`Has Source`,`Has Custom`)

  data_view
  }
  else{
    paste0('No results found for <a href="',discoURL(),'" target="_blank">',discoURL(),'</a>')
  }
}
# allow html hyperlinks
, sanitize.text.function = function(x) x)
```

### About

The Metadata Score is calculate by examining dataset `description`, `category`, `keywords/tags`, `source/attribution` and `custom metadata` and assigning a zero if there is no value or 1 if there is any value. The weight of each attribute can also be set from 0 to 1 via the input controls. Then the sum of the 5 attribute scores is divided by the sum of all the weights and finally multiplied by 100.

Data portals powered by [Socrata](https://socrata.com/).

```{r print_disco_url, eval=F}
# helpful for debugging
renderTable({
  url <- discoURL()
  paste0('<a href="',url,'" target="_blank">',url,'</a>')
}
# allow html hyperlinks
, sanitize.text.function = function(x) x)
```
