---
title: "Top Data by State"
author: "Alicia Brown"
output: 
  flexdashboard::flex_dashboard:
    orientation: rows
    social: menu
    source_code: embed
    favicon: favicon.png
runtime: shiny
---
```{r global, include=FALSE}
library(dplyr)
library(flexdashboard)
library(ggplot2)
library(httr)
library(jsonlite)
library(readr)
library(stringr)
library(tidyr)
library(tidytext)
library(wordcloud)
require(wesanderson)
```

```{r}
discoURL <- reactive({
  discovery_url <- paste0("http://api.us.socrata.com/api/catalog/v1?published=true"
                          ,"&only=",input$asset_type
                          ,"&domains=",input$domain
                          ,"&search_context=",input$domain
                          ,"&limit=",input$n_datasets
                          ,"&order=",input$rank)
  discovery_url
})
```

```{r}
# Reactive data available for all shiny modules
discoData <- reactive({
  
  datasets <- fromJSON(discoURL(),simplifyDataFrame = TRUE)
  resource <- datasets$results$resource
  if (is.null(resource) == FALSE){
  
    raw_data <- resource %>%
      select(id,name,description,download_count,updatedAt,createdAt)
    
    # page view stats
    page_views <- resource$page_views
    data <- cbind(raw_data, page_views)
    
    # domain category
    classification <- datasets$results$classification
    data <- cbind(data, classification$domain_category) %>%
      rename(category = `classification$domain_category`)
    
    # set selected rank variable for plots to adjust y variable easily
    if(input$rank == 'page_views_last_month'){
     data <- data %>%
       mutate(measure_value = page_views_last_month)
    }
    if(input$rank == 'page_views_last_week'){
      data <- data %>%
       mutate(measure_value = page_views_last_week)
    }
    if(input$rank == 'page_views_total'){
      data <- data %>%
       mutate(measure_value = page_views_total)
    }
    
    # capitalize x variables and replace NAs with 0 in download_count
    data <- data %>%
      mutate(download_count = replace_na(download_count, 0)) %>%
      rename(Dataset = name,
             Category = category)
    
    data
  }else{
    NULL
  }
})
```

Inputs {.sidebar data-width=300}
-----------------------------------------------------------------------

Discover top data by:

```{r}
# Socrata State Domains
selectInput("domain", label = "State Domain:",
            choices = list(
                        `California Controller's Office` = "bythenumbers.sco.ca.gov",
                        `California Treasurer's Office` = "data.debtwatch.treasurer.ca.gov",
                        `Colorado` = "data.colorado.gov",
                        `Connecticut` = "data.ct.gov",
                        `Hawaii` = "data.hawaii.gov",
                        `Iowa` = "data.iowa.gov",
                        `Maine` = "data.maine.gov",
                        `Maryland` = "data.maryland.gov",
                        `Massachusetts Comptroller` = "cthru.data.socrata.com",
                        `Michigan` = "data.michigan.gov",
                        `New Jersey` = "data.nj.gov",
                        `New York` = "data.ny.gov",
                        `Oregon` = "data.oregon.gov",
                        `Pennsylvania` = "data.pa.gov",
                        `Texas` = "data.texas.gov",
                        `Utah` = "opendata.utah.gov",
                        `Vermont` = "data.vermont.gov",
                        `Washington` = "data.wa.gov"
                        ), 
            selected = "data.colorado.gov")
selectInput("asset_type", label = "Type:",
            choices = c("dataset","chart","view","map","story","datalens"), selected="dataset")
selectInput("n_datasets", label = "Number:",
            choices = c(5, 10, 15, 20, 25), selected = 10)
# download_count not available to pass to order parameter
selectInput("rank", label = "Rank by:",
            choices = list(`Views Last Month` = "page_views_last_month",
           `Views Last Week` = "page_views_last_week",
           `Total Views` = "page_views_total"), selected = "page_views_last_month")
renderText("Powered by https://socratadiscovery.docs.apiary.io.")
```

Row
-----------------------------------------------------------------------

### Top Public Data
    
```{r fig.width=18, fig.height=8}
renderPlot({
  
  data_view <- discoData()
  
  if(length(data_view) > 0){
    # set sort order (todo: check out forcats package)
    data_view$Dataset <- factor(data_view$Dataset, levels=unique(data_view$Dataset[order(data_view$measure_value)]))
    
    # create Title of all currently selected variables in friendly format and make plot text more easily readable
    ggplot(data_view) +
    geom_bar(mapping = aes(x = Dataset, y = measure_value), na.rm = TRUE, stat = "identity") +
    coord_flip() +
    labs(x = '', y = '', 
         title=paste0('Top ', input$n_datasets,' ', input$asset_type,'s for ', input$domain,' by ',str_to_title(str_replace_all(input$rank,'_',' ')))
         ) +
    theme(text = element_text(size=15), plot.title = element_text(size=15, hjust=0.5))
  }
})
```

Row {.tabset}
-----------------------------------------------------------------------

### Metadata Description Cloud
    
```{r fig.width=18, fig.height=10}
pal <- wes_palette("Darjeeling1")
		
renderPlot({
  
  data_view <- discoData()
  
  if(length(data_view) > 0){
  
    dataset_descriptions <- data_view %>%
      unnest_tokens(word, description) %>%
      select(word) %>%
      anti_join(stop_words) %>%
      count(word) %>%
      with(wordcloud(word, n, max.words = 100, colors=pal,scale=c(3.0,0.5), rot.per=0.1))
  }
})
```

### Category Metadata

```{r fig.width=18, fig.height=10}
renderPlot({
  data_view <- discoData()

  if(length(data_view) > 0){
    data_view %>%
    group_by(Category) %>%
    summarize(download_count = sum(download_count),
              page_views_last_month = sum(page_views_last_month),
              page_views_last_week = sum(page_views_last_week),
              page_views_total = sum(page_views_total))
    # set selected rank variable for plots to adjust y variable easily
    if(input$rank == 'page_views_last_month'){
     data_view <- data_view %>%
       mutate(measure_value = page_views_last_month)
    }
    if(input$rank == 'page_views_last_week'){
      data_view <- data_view %>%
       mutate(measure_value = page_views_last_week)
    }
    if(input$rank == 'page_views_total'){
      data_view <- data_view %>%
       mutate(measure_value = page_views_total)
    }
  # set sort order and also had to invoke reorder for x aesthetic to sort decreasing
  data_view$Category <- factor(data_view$Category, levels=unique(data_view$Category[order(data_view$measure_value)]))
  # create Title of all currently selected variables in friendly format and make plot text more easily readable
  ggplot(data_view) +
    geom_bar(mapping = aes(x = reorder(Category, -measure_value), y = measure_value), na.rm = TRUE, stat = "identity") +
    labs(x = '', y = '', title=paste0('Category Metadata for Top ',input$n_datasets,' ',input$asset_type,'s by ',str_to_title(str_replace_all(input$rank,'_',' ')))) +
    theme(text = element_text(size=15), axis.text.x = element_text(angle=45, hjust=1),plot.title = element_text(size=15, hjust=0.5))
  }
})
```
   
### All Statistics

```{r}
renderTable({
  data_view <- discoData() 
  
  if(length(data_view) > 0){
    # format numbers with commas so easier to read in table view
    data_view  %>%
      mutate(download_count = format(download_count, big.mark=","),
             page_views_last_month = format(page_views_last_month, big.mark=","),
             page_views_last_week = format(page_views_last_week, big.mark=","),
             page_views_total = format(page_views_total, big.mark=",")) %>%
      rename(`Views Last Month` = page_views_last_month,
             `Views Last Week` = page_views_last_week,
             `Views Total` = page_views_total,
             `Downloads Total` = download_count) %>%
      select(Dataset,`Views Last Month`,`Views Last Week`,`Views Total`,`Downloads Total`)
    data_view
  }
})
```

### About

The Discovery API returns data found on Socrata client domains. Without any authentication, only public assets will be returned. Credentialed users may query their private assets as well. For more information on the API, check out https://socratadiscovery.docs.apiary.io.

The available state data portals are powered by [Socrata](https://socrata.com/).